Hyper-linked Communications: WebRTC enabled asynchronous collaboration

### 2016-06-09 curriculo/cv.tex
[RW] until qr code detection (reversed)

\entry
    {08 2015\\(standby)}
    {Systems Development}
    {Bullray-CIT}
    {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users through \underline{ffmpeg}. I tried \underline{FFServer}, \underline{nginx with rtmp module} and \underline{Wowza Streaming Server}, the final decision was \underline{Kurento Media Server} with an \underline{RTP} endpoint for receiving video from network cameras and a \underline{WebRTC} endpoint for delivering the video to web browsers. }}

\entry
    {03 2015\\(ongoing)}
    {Master Thesis}
    {IST}
    {\emph{Development of an interactive multi user video chat that supports recording, content overlay and a collaborative text editor. Used \underline{Kurento Media Server} for mixing streams into a single one and detecting QR codes. Experience with \underline{ot.js} for operation transformations (collaborative component). Other libraries used, \underline{jquery.qrcode}, \underline{vis.js}, \underline{bootstrap-typeahead}, \underline{codemirror.js} and \underline{adapter.js} }. Little experience with \underline{strophe.js} for communication between web browsers and \underline{XMPP} servers. }

@@

\section{interests}
    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging, Karate and Yoga.

>>>>

\entry
    {08 2015\\(2 months)}
    {Systems Development}
    {Bullray-CIT}
    {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users. I tried \underline{FFServer}, \underline{nginx with rtmp module} and \underline{Wowza Streaming Server}, the final decision was \underline{Kurento Media Server} with \underline{WebRTC}. }}

\entry
    {03 2015\\(1 year)}
    {Master Thesis \\\href{https://citysdk.tagus.ist.utl.pt:9443/}{try it}}
    {IST}
    {\emph{Development of an interactive multi user video chat that supports recording, content overlay and a collaborative text editor. Used \underline{Kurento Media Server} for mixing streams into a single one and detecting QR codes. Experience with \underline{ot.js} for operation transformations (collaborative component). Other libraries used, \underline{jquery.qrcode}, \underline{vis.js}, \underline{bootstrap-typeahead}, \underline{codemirror.js} and \underline{adapter.js} }. Little experience with \underline{strophe.js} for communication between web browsers and \underline{XMPP} servers. }

@@

\section{interests}
    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging, Kayaking, Karate and Yoga.

### 2016-06-09 tese2/16_implementation_content.tex
[RW] until qr code detection (reversed)

	\ac{KMS} lets registering event handlers for \ac{QR} code detection. The component that detects \ac{QR} codes on \ac{KMS} is called periodically and fires the handlers with the decoded content. This mechanism does not detect if the \ac{QR} code enters or leaves the screen. We had to implement our own mechanism for detecting those events. Each user session in the server maintains a map with the contents that are present on the screen. In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method. If the hash is not present in the map, it means the \ac{QR} code was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real time. If the same \ac{QR} code is not detected after a time bigger then two periods we can conclude that the \ac{QR} code left the screen and we add the correspondent content into our database. Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \ac{QR} code leaving detection.

@@

	In order to allow users to visualize past and present communications, we express the user's position as the time offset between its navigated time and the current time.

	In our solution, content is represented as simple text that can contain \ac{HTML}, \ac{CSS} or even \emph{JavaScript}. This content is displayed on defined intervals of time. Multiple contents can be displayed at the same time. In order to achieve that, we define layers above the video with the same size. Each layer is associated to the content.	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of the content instead of all of it, which would be very inefficient. The content to display for each user depends on the user's position on the timeline. This position is given by an offset between the current time and the navigated time. If the user is watching the content in real time this offset is always zero. If the user is reproducing past communications in normal playback rate, this offset is constant and, as a consequence, there is need to re-synchronize this value. If, for instance, we had to implement communication reproduction with a faster or slower playback rate, we would need to take into account the fact of this time offset not being constant.

	Each content is divided into two components, the \emph{start} and the \emph{end} which are sent to users. The \emph{start} component contains a time stamp, the content identification number and the content itself in form of text. The \emph{end} component only contains the time stamp and the content identification number.

@@

	The content to return is given by the union of the two following content subsets:
	\begin{itemize}
		\item Content that should be currently on display.
		\item A subset of contents whose starting time is immediately following the user's time.
	\end{itemize}

@@

	Our solution is flexible on what kind of interactions are possible to the users in real time, but allowing users to write \emph{JavaScript} that is executed on the other users browser would attackers to misuse their resources and access critical information. We could solve this problem easily by escaping any \emph{script} tag present on the content, but we would sacrifice the kind of interactions that are possible. By not allowing \emph{JavaScript} we would need to implement a subset of conceivable actions a priori and fire them when a type of message is received. We decided to ignore the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because, in order to offer the same interactions, we would have to do an exhaustive functional requirements gathering and, in our prototype, our goal is to explore users behavior when interacting with our system.

@@

		In order to show all streams to all users, each conference participant could receive individual streams for each user that share video. In overall, if there were $n$ active participants in the conference room, the number of streams per user would be $n$ ($1$ for sending video and $n-1$ for receiving) and, consequently, the number of streams in that conference room, as observed in Figure \ref{fig:xcomposite}, would be $n^2$.

@@

		By mixing all streams into a single one, users can receive a single stream with the representation of all active users. With this approach, the total number of streams within a conference room can be, at most, $2n$ (each user has $2$ streams, one to send and another to receive) as observed in Figure \ref{fig:ecomposite}.

>>>>

	\ac{KMS} allows registering event handlers for \ac{QR} code detection. The component that detects \ac{QR} codes on \ac{KMS} is called periodically and fires the handlers with the decoded content. Although the \ac{QR} code detection period is defined by \ac{KMS} we observed to be one second. This mechanism does not detect if the \ac{QR} code enters or leaves the screen. We had to implement our own mechanism for detecting those events. Each user session in the server maintains a map with the contents that are present on the screen, which can be represented by multiple \ac{QR} codes simultaneously. In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method. If the hash is not present in the map, it means the \ac{QR} code was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real time. If the same \ac{QR} code is not detected after a time bigger then two periods we can conclude that the \ac{QR} code left the screen and we add the correspondent content into our database. Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \ac{QR} code leaving detection.

@@

	In order to allow users to visualize past and present communications, we express the user's time position as the time offset between its reproducing time and the current time.

	In our solution, content is represented as simple text that can contain \ac{HTML}, \ac{CSS} or even \emph{JavaScript}. This content is displayed on defined intervals of time. Multiple contents can be displayed at the same time. In order to achieve that, we define layers above the video with the same size. Each layer is associated to the content.	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of the content instead of all of it, which would be very inefficient. The content to display for each user depends on the user's position on the timeline. This position is given by an offset between the current time and the reproducing time. If the user is watching the content in real time this offset is always zero. If the user is reproducing past communications in normal playback rate, this offset is constant and, as a consequence, there is need to re-synchronize this value. If, for instance, we had to implement communication reproduction with a faster or slower playback rate, we would need to take into account the fact of this time offset not being constant.

	Each content, that users receive from our application server, is divided into two components, the \emph{start} and the \emph{end}. The \emph{start} component contains a time stamp, the content identification number and the content itself in form of text. The \emph{end} component only contains the time stamp and the content identification number.

@@

	The content to return is given by the union of the two following content subsets:
	\begin{itemize}
		\item Content that should be currently on display.
		\item A subset of contents whose starting time is immediately after the user's time.
	\end{itemize}

@@

	Our solution is flexible on what kind of interactions are possible to the users in real time, but allowing users to write \emph{JavaScript} that is executed on the other users' browser would allow attackers to misuse their resources and access critical information. We could solve this problem easily by escaping any \emph{script} tag present on the content, but we would sacrifice the kind of interactions that are possible. By not allowing \emph{JavaScript} we would need to implement a subset of conceivable actions a priori and fire them when a type of message is received. We decided to postpone, to future work, the correction of the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because, in order to offer the same interactions, we would have to do an exhaustive functional requirements gathering and, in our prototype, our goal is to explore users behavior when interacting with our system.

@@

		In order to show all streams to all users, each conference participant could receive individual streams for each user that share video. If there are $n$ active participants in the conference room, the number of streams per user is $n$ ($1$ for sending video and $n-1$ for receiving) and, consequently, the number of streams in that conference room, as observed in Figure \ref{fig:xcomposite}, is $n^2$.

@@

		By mixing all streams into a single one, users can receive a single stream with the image (in a grid) and audio of all active users. With this approach, the total number of streams within a conference room can be, at most, $2n$ (each user has $2$ streams, one to send and another to receive) as observed in Figure \ref{fig:ecomposite}.

### 2016-06-09 tese2/evaluation.tex
[RW] until qr code detection (reversed)

  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.

@@

      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second: CPU, physical memory information relative to each running process and network usage relative to each interface.

@@

      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room. The second phase consists on the users leaving the conference room. Each event, joining and leaving, occurs with intervals of one minute in total of thirteen minutes (780 seconds).

@@

If we consider all streams equal and an incoming stream uses $x$kbps, with $n$ incoming streams the rate of data received at \ac{KMS} is $nx$ and the expected rate of data transfered to \emph{Kurento Repo} using localhost is $(n+1)x$. Due to the data being transfered also from \emph{Kurento Repo} to \emph{MongoDB} via localhost, the expected total rate of data transfered through localhost is $2((n+1)x) = (2n+2)x$, which explains the reason for the average amount of data transfered through localhost being more than twice the average amount of data received from users.

With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \ac{KMS} should be installed on the same machine, as depicted on Figure \ref{fig:scale2}, because the loopback interface can handle bigger transfer rates than the remaining network interfaces. Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \ac{KMS} and \emph{Kurento Repository} on the same machine.

@@

In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.

@@

In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics. In general the results reveal a significant difference between observed values and optimal values, which can be explained by the user's lack of experience with systems such as the one we have created.

>>>>

  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's user interface usability.

@@

      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects to a text file with a periodicity of one second: CPU, RAM usage relative to each running process and network usage relative to each interface.

@@

      The performance test scenario that we have defined consists of two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room. The second phase consists on the users leaving the conference room. Each event, joining and leaving, occurs with intervals of one minute in total of 13 minutes (780 seconds).

@@

If we consider all streams equal and an incoming stream uses $x$kbps, with $n$ incoming streams the rate of data received at \ac{KMS} is $nx$ and the expected rate of data transfered to \emph{Kurento Repo} using localhost is $(n+1)x$. Due to the data being transfered also from \emph{Kurento Repo} to \emph{MongoDB} via localhost, the expected total rate of data transfered through localhost is $2((n+1)x) = (2n+2)x$, which explains the reason for the average amount of data transfered through localhost being more than twice the average amount of data received from users. With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \ac{KMS} should be installed on the same machine, as depicted on Figure \ref{fig:scale2}, because the loopback interface can handle bigger transfer rates than the remaining network interfaces. Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \ac{KMS} and \emph{Kurento Repository} on the same machine.

@@

%In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.

@@

In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics. In general the results reveal a significant difference between observed values and optimal values, which can be explained by the user's lack of experience with with the tools that we provided.
